---
title: "ESG (FP1)"
author: "Orianna Wang"
date: "9/14/2022"
output: html_document
---

## Research questions

How water scarcity may make an impact on a corporate’s financial performance (e.g. bonds)? 

How does climate change impact agriculture and food prices?

Will climate change affect global trade (eg. import of necessities)? 






## Data 1: IFC Investment By Industry - Annual Summary

### Data source:
Provided by World Bank Group. IFC Industries (Sectors) are in charge of processing IFC transactions. 

### Data description:
The data shows the amount of money in millions of dollars the international finance cooperation investment on the commercial bases to different industries.

### Data limitation
Data As Of	06/30/2020, so it is recent. It contains investment in 10 industries, basically covering all of the industries that we need to examine investment and finance change with respect to time. 

### Variable names:
fiscal year: The fiscal year begins on July 1st of the previous year and runs through June 30th of the designated year.
Agribusiness and Forestry: investment in the industry of Agribusiness and Forestry in millions of USD.
Natural Resources:  investment in the industry of Natural Resources in millions of USD.

### Data visualization:

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(leaflet)

IFC <- read_csv("IFC/IFC_Investment_By_Industry_-_Annual_Summary.csv")
```
```{r}
IFC <- IFC[,1:6]
IFC
```

```{r}
IFC %>% 
  ggplot() +
  geom_line(aes(x = `Fiscal Year`,
                y = `Agribusiness and Forestry`),
            color = "red") +
  geom_line(aes(x = `Fiscal Year`,
                y = `Telecommunications and Information Technologies`),  
            color = "blue") +
  geom_line(aes(x = `Fiscal Year`,
                y = `Manufacturing`),
            color ="yellow") +
  geom_line(aes(x = `Fiscal Year`,
                y = `Infrastructure`)) +
  geom_line(aes(x = `Fiscal Year`,
                y = `Natural Resources`),
            color="grey") +
  labs(x = "Year",
       y = "Amount of investment (million USD)")
```



## Data 2: Agricultural land

### Data source
food and Agriculture Organization, electronic files and web site.

### Data description
Agricultural land refers to the share of land area that is arable, under permanent crops, and under permanent pastures. 

### Data limitations
the data is updated to 2018, lack recent data, but the data is not too old to lack analysis value. It contains agriculture land as a percentage of land area for each country, which is what I need. 

### Variable names
Country Name: the name of each country
1961: the agriculture land as a percentage of land area for each country in 1961. 
...same for each year

### Data visualization

```{r}
agriculture <- read_csv("arguculture/Argiculture.csv")
```
```{r}
agr_clean <- agriculture %>% 
  tail(-4) %>% 
  select(-c(...3, ...4, ...5)) %>% 
  pivot_longer(cols =starts_with("..."),
               names_to = "year",
               values_to = "Value")

agr_clean %>% 
  cbind(Year = as.numeric(substring(agr_clean$year, 4))) %>% 
  mutate(Year = Year + 1955) %>% 
  select(`Data Source`, Year, Value) %>% 
  group_by(Year) %>% 
  summarise(world_mean = mean(Value, na.rm = TRUE)) %>% 
  ggplot(aes(x = Year,
             y = world_mean),
         show.legend = FALSE) +
  geom_line()
  
```


Indicate some variable or variables that can connect at least two of the dataset in a join. You do not need to actually do the join, just figure out which variables would be helpful.


### Connection:
we can connect the investment in Agribusiness and Forestry change and the global average agriculture land use change to see the trend of agriculture with respect to time. 

